---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# crew

<!--[![CRAN](https://www.r-pkg.org/badges/version/crew)](https://CRAN.R-project.org/package=crew)-->
[![status](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![check](https://github.com/ropensci/crew/workflows/check/badge.svg)](https://github.com/ropensci/crew/actions?query=workflow%3Acheck)
[![codecov](https://codecov.io/gh/ropensci/crew/branch/main/graph/badge.svg?token=3T5DlLwUVl)](https://app.codecov.io/gh/ropensci/crew)
[![lint](https://github.com/ropensci/crew/workflows/lint/badge.svg)](https://github.com/ropensci/crew/actions?query=workflow%3Alint)

The [`R6`](https://r6.r-lib.org) classes of `crew` establish a standardized user interface to high-performance computing technologies. Unlike its closely related [`future`](https://future.futureverse.org/) package, `crew` prioritizes centralized scheduling, semi-persistent (reusable) workers, and user-driven customization. The primary goal is to help pipeline tools such as such [`targets`](https://docs.ropensci.org/targets/) efficiently orchestrate tasks without having to support individual low-level interfaces to specific high-performance computing platforms or cloud services.

## Installation

```{r, eval = FALSE}
remotes::install_github("ropensci/crew")
```

## Code of Conduct

Please note that the `crew` project is released with a [Contributor Code of Conduct](https://github.com/wlandau/crew/blob/main/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.

## Citation

```{r, warning = FALSE}
citation("crew")
```
