---
output: github_document
---

```{r, include = FALSE}
suppressPackageStartupMessages({
  library(crew)
  library(future)
  library(future.callr)
  library(purrr)
})
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# crew

<!--[![CRAN](https://www.r-pkg.org/badges/version/crew)](https://CRAN.R-project.org/package=crew)-->
[![status](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![check](https://github.com/wlandau/crew/workflows/check/badge.svg)](https://github.com/wlandau/crew/actions?query=workflow%3Acheck)
[![codecov](https://codecov.io/gh/wlandau/crew/branch/main/graph/badge.svg?token=3T5DlLwUVl)](https://app.codecov.io/gh/wlandau/crew)
[![lint](https://github.com/wlandau/crew/workflows/lint/badge.svg)](https://github.com/wlandau/crew/actions?query=workflow%3Alint)

A task queue is a central hub for orchestrating
  computationally demanding workloads dynamically and asynchronously.
  The `crew` package supports a variety of task queues with
  a common interface and different types of workers for different
  backend technologies, similar to the way the
  [`future`](https://future.futureverse.org/) package
  provides a unified R interface for single jobs and
  individual map-reduce calls. The `crew` package borrows
  the task queue design from the [`callr` package vignettes](https://github.com/r-lib/callr/blob/811a02f604de2cf03264f6b35ce9ec8a412f2581/vignettes/Task-queue-with-callr.Rmd) and
  <https://www.tidyverse.org/blog/2019/09/callr-task-q/>,
  enhances the core feature set (for example, crash detection),
  and establishes an extensible [`R6`](https://r6.r-lib.org/)
  class system to add more worker types
  as development continues (e.g. jobs on Amazon and Google cloud
  computing platforms). Multi-queue structures, such as the
  already-supported '[`future`](https://future.futureverse.org/)-based task queue, increase
  efficiency in heavily parallel [cluster](https://future.batchtools.futureverse.org/) workloads
  with an inner local task queue to asynchronicity manage
  task submission, monitoring, and collection.
  
## Similar work

* [`targets`](https://docs.ropensci.org/targets/): a Make-like pipeline tool for R. Event loops in [`targets`](https://docs.ropensci.org/targets/) invoke packages [`clustermq`](https://mschubert.github.io/clustermq/) and [`future`](https://future.futureverse.org/) to run tasks concurrently and asynchronously. [`targets`](https://docs.ropensci.org/targets/) would be able to support different scheduling techniques and run on more platforms if it incorporated a formal task queue behind the scenes. That task queue is probably going to be `crew` or [`rrq`](https://mrc-ide.github.io/rrq/).
* [`callr`](https://github.com/r-lib/callr): the previous [task queue  vignette](https://github.com/r-lib/callr/blob/811a02f604de2cf03264f6b35ce9ec8a412f2581/vignettes/Task-queue-with-callr.Rmd), as well as the blog post at <https://www.tidyverse.org/blog/2019/09/callr-task-q/>, show how to create an efficient multi-process task queue using [`callr::r_session`](https://callr.r-lib.org/reference/r_session.html). The goal of `crew` is to eventually extend this idea to different worker types, such as [Amazon Web Services (AWS) Batch](https://aws.amazon.com/batch/) jobs.
* [`rrq`](https://mrc-ide.github.io/rrq/): an R-focused task queue built on Redis, successor of [`rrqueue`](http://traitecoevo.github.io/rrqueue/). It is possible that [`rrq`](https://mrc-ide.github.io/rrq/) makes `crew` irrelevant.
* [`rrqueue`](http://traitecoevo.github.io/rrqueue/): an R-focused task queue built on Redis, no longer maintained.
* [`future`](https://future.futureverse.org/): a unified interface for asynchronous evaluation of single tasks and map-reduce calls on a wide variety of backend technologies. One of the backends of `crew`.
* [`clustermq`](https://mschubert.github.io/clustermq/): an R layer on top of [ZeroMQ](https://zeromq.org/) for submitting tasks as array jobs on traditional clusters.
* [`mirai`](https://github.com/shikokuchuo/mirai): an R framework for asynchronous tasks built on [NNG](https://nng.nanomsg.org).
* [`promises`](https://rstudio.github.io/promises/): minimally-invasive asynchronous programming for a small number of tasks within Shiny apps.
* [`later`](https://r-lib.github.io/later/): delayed evaluation of synchronous tasks.
* [High-performance computing CRAN task view](	https://CRAN.R-project.org/view=HighPerformanceComputing).

## Installation

```{r, eval = FALSE}
remotes::install_github("wlandau/crew")
```

## Documentation

Please see <https://wlandau.github.io/crew/articles/usage.html> for a tutorial.

## Future development

`crew`, [`rrq`](https://mrc-ide.github.io/rrq/), or something like it will serve as a backend to get [`targets`](https://docs.ropensci.org/targets/) running distributed workers on the cloud (AWS Batch, AWS Fargate, Google Cloud Run, etc.). If `crew` fills this role, new [task queues](https://wlandau.github.io/crew/reference/index.html#queues) and [data stores](https://wlandau.github.io/crew/reference/index.html#stores) will be developed for cloud platforms.

## Thanks

The `crew` package incorporates insightful ideas from the following people.

* [Gábor Csárdi](https://github.com/gaborcsardi/). Gábor created the [`callr`](https://github.com/r-lib/callr) R package and wrote instructional code for creating task queues using [`callr::r_session`](https://callr.r-lib.org/reference/r_session.html) (links below). `crew` borrows and modifies the [vignette code](https://github.com/r-lib/callr/blob/811a02f604de2cf03264f6b35ce9ec8a412f2581/vignettes/taskq.R) under the MIT license. See also the `crew` package `NOTICE` file.
  * Blog post: <https://www.tidyverse.org/blog/2019/09/callr-task-q/>
  * `callr` vignette code: <https://github.com/r-lib/callr/blob/811a02f604de2cf03264f6b35ce9ec8a412f2581/vignettes/taskq.R>
* [Kirill Müller](https://github.com/krlmlr/). The [`workers`](https://github.com/wlandau/workers) prototype was entirely his vision, and `crew` would not exist without it. `crew` reflects this and many other insights from Kirill about orchestration models for R processes.
* [Henrik Bengtsson](https://github.com/HenrikBengtsson/). Henrik's [`future`](https://github.com/HenrikBengtsson/future/) package ecosystem demonstrates the incredible power of a consistent R interface on top of a varying collection of high-performance computing technologies.
* [Michael Schubert](https://github.com/mschubert/). Michael's [`clustermq`](https://mschubert.github.io/clustermq/) package supports efficient high-performance computing on traditional clusters, and it demonstrates the value of a central `R6` object to manage an entire collection of persistent workers.
* [David Kretch](https://github.com/davidkretch). The [`paws`](https://github.com/paws-r/paws) R package is a powerful interface to Amazon Web Services, and the documentation clearly communicates the capabilities and limitations of AWS to R users.
* [Adam Banker](https://github.com/adambanker), co-authored [`paws`](https://github.com/paws-r/paws) with [David Kretch](https://github.com/davidkretch).
* [David Neuzerling](https://github.com/mdneuzerling). David's [`lambdr`](https://github.com/mdneuzerling/lambdr/) package establishes a helpful pattern to submit and collect AWS Lambda jobs from R.
* [Mark Edmondson](https://github.com/MarkEdmondson1234/). Mark maintains several R packages to interface with Google Cloud Platform such as [`googleCloudStorageR`](https://github.com/cloudyr/googleCloudStorageR) and [`googleCloudRunner`](https://github.com/MarkEdmondson1234/googleCloudRunner), and he [started the conversation](https://github.com/ropensci/targets/issues/720) around helping [`targets`](https://github.com/ropensci/targets) submit jobs to Google Cloud Run.

## Code of Conduct

Please note that the `crew` project is released with a [Contributor Code of Conduct](https://github.com/wlandau/crew/blob/main/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.

## Citation

```{r, warning = FALSE, comment = ""}
citation("crew")
```
