---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# crew

<!--[![CRAN](https://www.r-pkg.org/badges/version/crew)](https://CRAN.R-project.org/package=crew)-->
[![status](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![check](https://github.com/ropensci/crew/workflows/check/badge.svg)](https://github.com/ropensci/crew/actions?query=workflow%3Acheck)
[![codecov](https://codecov.io/gh/ropensci/crew/branch/main/graph/badge.svg?token=3T5DlLwUVl)](https://app.codecov.io/gh/ropensci/crew)
[![lint](https://github.com/ropensci/crew/workflows/lint/badge.svg)](https://github.com/ropensci/crew/actions?query=workflow%3Alint)

The user interface in `crew` is an abstraction for high-performance computing technologies. Like the `future` R package, `crew` provides a backend-agnostic standard. But unlike `future`, `crew` prioritizes centralized scheduling, semi-persistent workers, and user-driven customization. The user supplies methods to initialize, dispatch, poll, and terminate a worker on a specific backend technology. The backend technology could be forked processes, an on-premises high-performance computing cluster,
a cloud computing service, or even another worker abstraction like `future`. Then, `crew` supplies an `R6` subclass to create objects that manage the actual workers. Every worker group has a centralized mutable cache and a centralized registry of all the workers in the group, both of which are available to the user-defined worker methods. This facilitates advanced capabilities such as worker affinities and batched polling. The end goal is to help pipeline tools such as such `targets` efficiently orchestrate tasks without having to support individual low-level interfaces to specific high-performance computing platforms or cloud computing services.

## Installation

```{r, eval = FALSE}
remotes::install_github("ropensci/crew")
```

## Code of Conduct

Please note that the `crew` project is released with a [Contributor Code of Conduct](https://github.com/wlandau/crew/blob/main/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.

## Citation

```{r, warning = FALSE}
citation("crew")
```
