---
output: github_document
---

```{r, include = FALSE}
suppressPackageStartupMessages({
  library(crew)
  library(future)
  library(future.callr)
  library(purrr)
})
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# crew

<!--[![CRAN](https://www.r-pkg.org/badges/version/crew)](https://CRAN.R-project.org/package=crew)-->
[![status](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![check](https://github.com/wlandau/crew/workflows/check/badge.svg)](https://github.com/wlandau/crew/actions?query=workflow%3Acheck)
[![codecov](https://codecov.io/gh/wlandau/crew/branch/main/graph/badge.svg?token=3T5DlLwUVl)](https://app.codecov.io/gh/wlandau/crew)
[![lint](https://github.com/wlandau/crew/workflows/lint/badge.svg)](https://github.com/wlandau/crew/actions?query=workflow%3Alint)

The [`R6`](https://r6.r-lib.org) classes of `crew` establish a standardized user interface to high-performance computing technologies. Unlike its closely related [`future`](https://future.futureverse.org/) package, `crew` prioritizes centralized scheduling, heterogeneous semi-persistent workers, and user-driven customization. The primary goal is to help pipeline tools such as such [`targets`](https://docs.ropensci.org/targets/) efficiently orchestrate tasks without having to support individual low-level interfaces to specific high-performance computing platforms or cloud services.

## Installation

```{r, eval = FALSE}
remotes::install_github("wlandau/crew")
```

## Usage

First, create a `crew` object.

```{r}
library(crew)
crew <- class_crew$new()
```

Optionally, supply the store and worker classes you plan to use. The store determines where the input and output data live for jobs (local file system, Amazon S3 bucket, Google Cloud bucket, etc.), and worker classes control where and how the workers run (locally in a different R session, Amazon Web Services, Google Cloud Platform, etc.).

```{r}
crew <- class_crew$new(
  store = class_store_local$new(),
  worker_classes = list(
    class_worker_callr,
    class_worker_future
  )
)
```

If you use `class_worker_future`, choose a plan.

```{r}
library(future)
library(future.callr)
plan(callr)
```

Create worker objects with the `recruit()` method.

```{r}
crew$recruit(workers = 2) # callr workers
```

Optionally, create worker objects of a specified class and set fields like timeout and tags.

```{r}
names(crew$worker_classes)
crew$recruit(
  workers = 2,
  class = "worker_future",
  tags = c("long_jobs_go_here", "my_future_workers")
)
```

No workers are running yet.

```{r}
library(purrr)
map_lgl(crew$workers, ~.x$up())
```

To start any or all workers, either use `launch()` or `send()`. The latter finds a suitable available worker, launches it if necessary, and then sends a job. Poll with `receivable()` and get the output with `receive()`. Tags let you control which kinds of workers get the jobs, and they are are supported and optional in all relevant crew methods.

```{r, warning = FALSE}
job <- function(seconds) {
  Sys.sleep(seconds)
  warning("This is a warning.")
  sprintf("Job ran in %s seconds.", seconds)
}

# The job runs on a worker of class "worker_future" in this case
# because those two workers are tagged with "long_jobs_go_here".
crew$send(fun = job, args = list(seconds = 4), tags = "long_jobs_go_here")

# Do other tasks while the job runs in the background.
do_other_tasks <- function () {
  start <- as.numeric(proc.time()["elapsed"])
  while (!crew$receivable()) {
    Sys.sleep(1)
    print(as.numeric(proc.time()["elapsed"]) - start)
  }
}

do_other_tasks()

# Get the output when ready.
output <- crew$receive()
```

The output contains the result of the job, as well as warnings, errors, etc.

```{r}
str(output)
```

Once launched, workers stay running until they time out or shut down.

```{r}
is_up <- map_lgl(crew$workers, ~.x$up())
is_up
```

Sure enough, the running worker has the tags of the previous job.

```{r}
name <- names(is_up[is_up])
crew$workers[[name]]$tags
```

`crew` tries to send jobs to workers that are already running. If all running workers are busy, new ones will launch automatically. Use `sendable()` to check if any workers can accept new work.

```{r, warning = FALSE}
crew$send(fun = job, args = list(seconds = 8), tags = "long_jobs_go_here")
crew$send(fun = job, args = list(seconds = 4)) # May run on any available worker.

do_other_tasks()

output_first <- crew$receive()

do_other_tasks()

output_second <- crew$receive()

str(output_first)

str(output_second)
```

To shut down one or more workers, use `shutdown()`. By default, only idle workers are shut down.

```{r}
crew$shutdown()
while (any(map_lgl(crew$workers, ~.x$up()))) {
  Sys.sleep(0.1)
}
```

To delete one or more worker objects from the crew, use `dismiss()`.

```{r}
crew$dismiss(tags = "long_jobs_go_here")
length(crew$workers) # Should be 2.
```

If eventually a worker gets stuck (not sendable, not receivable, not running) then the `stuck()` method of the worker will return `TRUE` and the `restart()` method will restart it. Use the `restart()` method of the crew object to restart one or more stuck workers.

## Nested crews

Large data science pipelines manage hundreds of distributed workers on different machines, from traditional clusters to cloud platforms by Amazon and Google. In these situations, orchestration itself is computationally demanding. Sending jobs, receiving job output, and detecting crashes may severely block the main process.

To overcome this bottleneck, `crew` package will eventually support nested crews. The outer crew will support a small number of `callr::r_bg()` workers running on the same machine as the main process. Each of these outer workers will run a crew of its own, and the inner crews will run hundreds of inner workers on distributed systems like AWS Batch. The outer workers will run a different event loop that forwards jobs from the outer crew to the inner crews and back.

## New backends

The `crew` package is arbitrarily extensible. It is designed to support multiple backend services (e.g. Amazon, Google, Azure) with minimal changes to the `R6` interface. For Amazon S3 storage for jobs, one can write a new subclass of `class_store` and write methods analogous to those of `class_store_local`. Likewise, for Amazon AWS Batch or Google Cloud Run workers, one can write new worker subclasses analogous to `class_worker_callr` and `class_worker_future`. All these classes are documented in help files, and a specification of required methods and fields is in the [specification vignette](https://wlandau.github.io/crew/articles/specification.html). The new subclasses will be supported in separate R packages that import `crew`, ideally one package per cloud platform.

## Thanks

The `crew` package incorporates insightful ideas from the following people.

* [Kirill Müller](https://github.com/krlmlr/). The [`workers`](https://github.com/wlandau/worker) prototype was entirely his vision, and `crew` would not exist without it. `crew` reflects this and many other insights from Kirill about orchestration models for R processes.
* [Henrik Bengtsson](https://github.com/HenrikBengtsson/). Henrik's [`future`](https://github.com/HenrikBengtsson/future/) package ecosystem demonstrates the incredible power of a consistent R interface on top of a varying collection of high-performance computing technologies.
* [Michael Schubert](https://github.com/mschubert/). Michael's [`clustermq`](https://mschubert.github.io/clustermq/) package supports efficient high-performance computing on traditional clusters, and it demonstrates the value of a central `R6` object to manage an entire collection of persistent workers.
* [David Kretch](https://github.com/davidkretch). The [`paws`](https://github.com/paws-r/paws) R package is a powerful interface to Amazon Web Services, and the documentation clearly communicates the capabilities and limitations of AWS to R users.
* [Adam Banker](https://github.com/adambanker), co-authored [`paws`](https://github.com/paws-r/paws) with [David Kretch](https://github.com/davidkretch).
* [David Neuzerling](https://github.com/mdneuzerling). David's [`lambdr`](https://github.com/mdneuzerling/lambdr/) package establishes a helpful pattern to submit and collect AWS Lambda jobs from R.
* [Mark Edmondson](https://github.com/MarkEdmondson1234/). Mark maintains several R packages to interface with Google Cloud Platform such as [`googleCloudStorageR`](https://github.com/cloudyr/googleCloudStorageR) and [`googleCloudRunner`](https://github.com/MarkEdmondson1234/googleCloudRunner), and he [started the conversation](https://github.com/ropensci/targets/issues/720) around helping [`targets`](https://github.com/ropensci/targets) submit jobs to Google Cloud Run.

## Code of Conduct

Please note that the `crew` project is released with a [Contributor Code of Conduct](https://github.com/wlandau/crew/blob/main/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.

## Citation

```{r, warning = FALSE}
citation("crew")
```
