---
title: "Controller groups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Controller groups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
set.seed(0)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  paged.print = FALSE,
  eval = FALSE
)
```

Each controller object only supports only one type of worker configuration which you set in advance. However, different controllers may have different types of workers, and `crew` supports controller groups to coordinate among these different worker types. With third-party launcher subclasses from other packages, this mechanism will allow you to e.g. send some tasks to GPU-capable or high-memory workers while other tasks go to low-spec workers.

We demonstrate with a controller of fully persistent workers which always stay running and a controller of semi-persistent workers which terminate after completing four tasks. We create controller objects with names.

```{r}
library(crew)
persistent <- crew_controller_local(name = "persistent")
transient <- crew_controller_local(name = "semi-persistent", tasks_max = 4L)
```

`crew` uses a different TCP port for each controller you run, so please do not create hundreds of controllers. Please see the subsection on ports in the [README](https://wlandau.github.io/crew/index.html).

We put these controller objects into a new controller group object.

```{r}
group <- crew_controller_group(persistent, transient)
```

This controller group has a global `connect()` method to initialize both controllers.

```{r}
group$start()
```

You can choose which worker pool to receive tasks.

```{r}
group$push(name = "my task", command = sqrt(4), controller = "semi-persistent")
```

The controller group also supports global methods for `wait()`, `pop()`, and `terminate()`. These methods operate on all controllers at once by default, but the `controllers` argument allows you to select a subset of controllers to act on. Below in `pop()` the `launcher` column of the output indicates which controller ran the task.

```{r}
group$wait(controllers = "semi-persistent")
group$pop()
#> # A tibble: 1 × 11
#>   name    command result seconds   seed error trace warnings launcher worker
#>   <chr>   <chr>   <list>   <dbl>  <int> <chr> <chr> <chr>    <chr>     <int>
#> 1 my task NA      <dbl>        0 6.30e8 NA    NA    NA       semi-pe…      1
#> # ℹ 1 more variable: instance <chr>
```

The controller group has a `summary()` method which aggregates the summaries of one or more controllers.

```{r}
group$summary()
#> # A tibble: 2 × 6
#>   controller      worker tasks seconds errors warnings
#>   <chr>            <int> <int>   <dbl>  <int>    <int>
#> 1 persistent           1     0       0      0        0
#> 2 semi-persistent      1     1       0      0        0
```

When you are finished, please call `terminate()` with no arguments to terminate all controllers in the controller group.

```{r}
group$terminate()
```
