---
title: "Multi-controllers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multi-controllers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
set.seed(0)
library(crew)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  paged.print = FALSE,
  eval = FALSE
)
```

Each controller object only supports only one type of worker configuration which you set in advance. However, different controllers may have different types of workers, and `crew` supports multi-controller functionality to coordinate among these different worker types. In future development, this mechanism will allow you to e.g. send some tasks to GPU-capable or high-memory workers while other tasks go to low-spec workers. But even now, this feature can come in handy.

We demonstrate with a controller of fully persistent workers which always stay running and a controller of fully transient workers which terminate as soon as their first task is complete. We create controller objects with names.

```{r}
persistent <- crew_controller_callr(name = "persistent")
transient <- crew_controller_callr(name = "transient", max_tasks = 1L)
```

We put these controller objects into a new multi-controller object.

```{r}
multi <- crew_multi_controller(persistent, transient)
```

This controller has a global `connect()` method to initialize both controllers.

```{r}
multi$connect()
```

You can choose which worker pool to receive tasks.

```{r}
multi$push(name = "my task", command = sqrt(4), controller = "transient")
```

The multi-controller also supports global methods for `wait()`, `pop()`, and `terminate()`. These methods operate on all controllers at once by default, but the `controllers` argument allows you to select a subset of controllers to act on.

```{r}
multi$wait(controllers = "transient")
multi$pop()
#> # A tibble: 1 × 7
#>   name    command result    seconds error traceback warnings
#>   <chr>   <chr>   <list>      <dbl> <chr> <chr>     <chr>
#> 1 my task sqrt(4) <dbl [1]>       0 NA    NA        NA
```

When you are finished, please call `terminate()` with no arguments to terminate all controllers in the multi-controller.

```{r}
multi$terminate()
```
