---
title: "Launcher plugins"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Launcher plugins}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library(crew)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# About

`crew` lets users write custom [launchers](https://wlandau.github.io/crew/reference/crew_class_launcher.html) for different types of workers that connect over the local network. The [`crew.cluster`](https://wlandau.github.io/crew.cluster/) package already has plugins for traditional high-performance computing schedulers ([SLURM](https://slurm.schedmd.com/), SGE, LSF, and PBS/TORQUE).

# 📣 Request for community contributions 📣

<span style="color: blue">**The launcher plugin framework aims to extend `crew` to modern platforms and services like [AWS Batch](https://aws.amazon.com/batch/), [Google Cloud Run](https://cloud.google.com/run/), [Kubernetes](https://kubernetes.io/), and beyond. At the time of writing, plugins for cloud computing do not yet exist. So if you have access to these services and know how to use them, please consider contributing a package with plugins of your own. The maintainer of `crew` would love to work with you!**</span>

# How it works

These [launcher](https://wlandau.github.io/crew/reference/crew_class_launcher.html) plugins need not become part of the `crew` package itself. You can write your plugin in a simple R script, or you write it in a custom R package that [depends on](https://r-pkgs.org/dependencies-in-practice.html) `crew`. Published packages with [launcher](https://wlandau.github.io/crew/reference/crew_class_launcher.html) plugins are powerful extensions that enhance `crew` for the entire open-source community. See [R Packages](https://r-pkgs.org/) by [Hadley Wickham](https://hadley.nz/) and [Jenny Bryan](https://jennybryan.org/) for how to write an R package.

# Scope

This vignette demonstrates how to write a `crew` launcher plugin. It assumes prior familiarity with [`R6` classes](https://r6.r-lib.org/articles/Introduction.html) and the computing platform of your plugin.

# Implementation

To create your own launcher plugin, write an [`R6`](https://r6.r-lib.org/articles/Introduction.html) subclass of [`crew_class_launcher`](https://wlandau.github.io/crew/reference/crew_class_launcher.html) with a [`launch_worker()`](https://wlandau.github.io/crew/reference/crew_class_launcher_local.html#method-launch-worker-) method analogous the one in the [local process launcher](https://wlandau.github.io/crew/reference/crew_class_launcher_local.html). `launch_worker()` must accept the same arguments as the [local process `launch_worker()` method](https://wlandau.github.io/crew/reference/crew_class_launcher_local.html#method-launch-worker-), generate a call to [`crew_worker()`](https://wlandau.github.io/crew/reference/crew_worker.html), and then submit a new job or process to run that call.

# Network

Each worker that launches must be able to dial into the client over the local network. The `host` argument of `crew_client()` provides the local IP address, and the `port` argument provides the TCP port. The controller helper function (see below) should expose arguments `host` and `port` in order to solve potential network problems like [this one](https://github.com/wlandau/crew.cluster/issues/1#issuecomment-1546024163).

By default, `host` is the local IP address. `crew` assumes the local network is secure. Please take the time to assess the network security risks of your computing environment. Use at your own risk.

# Termination

We recommend you implement an optional [`terminate_worker()`](https://wlandau.github.io/crew/reference/crew_class_launcher_local.html#method-terminate-worker-) method. Although `mirai` has its own way of terminating workers, it only works if the worker already connected, and it cannot reach workers that fail to connect and hang in a crashed state. An optional [`terminate_worker()`](https://wlandau.github.io/crew/reference/crew_class_launcher_local.html#method-terminate-worker-) method in your `crew` launcher plugin is extra assurance that these workers will exit.

If you implement a custom [`terminate_worker()`](https://wlandau.github.io/crew/reference/crew_class_launcher_local.html#method-terminate-worker-) method, it must not throw an error (and should not throw a warning or message) if the worker is already terminated. In addition, it must accept a handle that identifies the worker, and this handle must be the return value of the previous call to [`launch_worker()`](https://wlandau.github.io/crew/reference/crew_class_launcher_local.html#method-launch-worker-). A handle can be any kind of R object: a process ID, a job name, an `R6` object returned by `callr::r_bg()`, etc.

# Example

The following is a custom custom launcher class whose workers are local R processes on Unix-like systems.^[See [`tests/launchers/test-launcher-system2.R`](https://github.com/wlandau/crew/blob/main/tests/launchers/test-launcher-system2.R) for an example launcher without a `termiante_worker()` method.]

```{r}
custom_launcher_class <- R6::R6Class(
  classname = "custom_launcher_class",
  inherit = crew::crew_class_launcher,
  public = list(
    launch_worker = function(call, name, launcher, worker, instance) {
      bin <- file.path(R.home("bin"), "R")
      processx::process$new(
        command = bin,
        args = c("-e", call),
        cleanup = FALSE
      )
    },
    terminate_worker = function(handle) {
      handle$kill()
    }
  )
)
```

Inside `launch_worker()`, the `processx::process$new(command = bin, args = c("-e", call))` line runs the [`crew_worker()`](https://wlandau.github.io/crew/reference/crew_worker.html) call in an external R process. This process runs in the background, connects back to `crew` and `mirai` over the local network, and accepts the tasks you push to the controller. `processx::process$new()` also returns a handle which the `terminate_worker()` method can use to force-terminate the process if appropriate. `mirai` has its own way to terminate workers, so a `terminate_worker()` method is not strictly required, but it is a useful safeguard in case a worker hangs in a crashed state before it establishes a connection.

Every `launch_worker()` method must accept arguments `call`, `launcher`, `worker`, and `instance`. The method does not actually need to use all these arguments, but they must be present in the function signature.

* `call`: text string with a call to  [`crew_worker()`](https://wlandau.github.io/crew/reference/crew_worker.html).
* `name`: text string with an informative name for the worker instance.
* `launcher`: text string with the name of the launcher.
* `worker`: positive integer index of the worker. Ranges from 1 to the maximum number of simultaneous workers configured for the controller.
* `instance`: text string with the instance of the worker in the launcher at the given worker index.


To see what the `call` object looks like, create a new launcher and run the `call()` method.

```{r, eval = TRUE, message = FALSE}
library(crew)
launcher <- crew_launcher_local()
launcher$call(
  socket = "ws://127.0.0.1:5000/3/aa9c59ea",
  launcher = "my_launcher",
  worker = 3L,
  instance = "aa9c59ea"
)
```

# Helper

It is useful to have a helper function that creates controllers with your custom launcher. It should:

1. Accept all the same arguments as [`crew_controller_local()`](https://wlandau.github.io/crew/reference/crew_controller_local.html).
2. Create a client object using [`crew_client()`](https://wlandau.github.io/crew/reference/crew_client.html).
3. Create a launcher object with the [`new()` method](https://wlandau.github.io/crew/reference/crew_class_launcher.html#method-crew_class_launcher-new) of your custom launcher class.
4. Create a new controller using [`crew_controller()`](https://wlandau.github.io/crew/reference/crew_controller.html).
5. Scan the controller for obvious errors using the [`validate()`](https://wlandau.github.io/crew/reference/crew_class_controller.html#method-crew_class_controller-validate) method of the controller.

Feel free to borrow from the [`crew_controller_local()` source code](https://github.com/wlandau/crew/blob/main/R/crew_controller_local.R). For packages, you can use the `@inheritParams` [`roxygen2`](https://roxygen2.r-lib.org/) tag to inherit the documentation of all the arguments instead of writing it by hand. You may want to adjust the default arguments based on the specifics of your platform, especially `seconds_launch` if workers take a long time to launch.

```{r}
#' @title Create a controller with the custom launcher.
#' @export
#' @description Create an `R6` object to submit tasks and
#'   launch workers.
#' @inheritParams crew::crew_controller_local
crew_controller_custom <- function(
  name = "custom controller name",
  workers = 1L,
  host = NULL,
  port = NULL,
  tls = crew::crew_tls(),
  seconds_interval = 0.5,
  seconds_timeout = 10,
  seconds_launch = 30,
  seconds_idle = Inf,
  seconds_wall = Inf,
  seconds_exit = 1,
  tasks_max = Inf,
  tasks_timers = 0L,
  reset_globals = TRUE,
  reset_packages = FALSE,
  reset_options = FALSE,
  garbage_collection = FALSE,
  launch_max = 5L
) {
  client <- crew::crew_client(
    name = name,
    workers = workers,
    host = host,
    port = port,
    tls = tls,
    seconds_interval = seconds_interval,
    seconds_timeout = seconds_timeout
  )
  launcher <- custom_launcher_class$new(
    name = name,
    seconds_interval = seconds_interval,
    seconds_launch = seconds_launch,
    seconds_idle = seconds_idle,
    seconds_wall = seconds_wall,
    seconds_exit = seconds_exit,
    tasks_max = tasks_max,
    tasks_timers = tasks_timers,
    reset_globals = reset_globals,
    reset_packages = reset_packages,
    reset_options = reset_options,
    garbage_collection = garbage_collection,
    launch_max = launch_max,
    tls = tls
  )
  controller <- crew::crew_controller(client = client, launcher = launcher)
  controller$validate()
  controller
}
```

# Informal testing

Before you begin testing, please begin monitoring local processes and remote jobs on your platform. In the case of the above `crew` launcher which only creates local processes, it is sufficient to start [`htop`](https://htop.dev/) and filter for R processes, or launch a new R session to monitor the process table from [`ps::ps()`](https://ps.r-lib.org/reference/ps.html). However, for more ambitious launchers that submit workers to e.g. [AWS Batch](https://aws.amazon.com/batch/), you may need to open the [CloudWatch](https://aws.amazon.com/cloudwatch/) dashboard, then view the AWS billing dashboard after testing.

When you are ready to begin testing, try out the example in the [README](https://wlandau.github.io/crew/index.html#usage), but use your your custom controller helper instead of [`crew_controller_local()`](https://wlandau.github.io/crew/reference/crew_controller_local.html). 

First, create and start a controller. You may wish to monitor local processes on your computer to make sure the `mirai` dispatcher starts.

```{r}
library(crew)
controller <- crew_controller_custom(workers = 2)
controller$start()
```

Try pushing a task that gets the local IP address and process ID of the worker instance.

```{r}
controller$push(
  name = "get worker IP address and process ID",
  command = paste(getip::getip(type = "local"), ps::ps_pid())
)
```

Wait for the task to complete and look at the result.

```{r}
controller$wait()
result <- controller$pop()
result$result[[1]]
#> [1] "192.168.0.2 27336"
```

Please use the result to verify that the task really ran on a worker as intended. The process ID above should agree with the one from the handle ([except on Windows](https://github.com/r-lib/processx/issues/364) because the actual R process may be different from the `R.exe` process created first). In addition, if the worker is running on a different computer, the worker IP address should be different than the local IP address. Since our custom launcher creates local processes, the IP addresses are the same in this case, but they should be different for a [SLURM](https://slurm.schedmd.com/) or [AWS Batch](https://aws.amazon.com/batch/) launcher. 

```{r}
getip::getip(type = "local")
#> "192.168.0.2"
controller$launcher$workers$handle[[1]]$get_pid()
#> [1] 27336
```

If you did not set any timeouts or task limits, the worker that ran the task should still be online. The other worker had no tasks, so it did not need to launch.

```{r}
controller$client$summary()
#> # A tibble: 2 × 6
#>   worker online instances assigned complete socket
#>    <int> <lgl>      <int>    <int>    <int> <chr>
#> 1      1 TRUE           1        1        1 ws://10.0.0.32:50258/1/571bcda7…
#> 2      2 FALSE          0        0        0 ws://10.0.0.32:50258/2/daf88d6e…
```

When you are done, terminate the controller. This terminates the `mirai` dispatcher process and the `crew` workers.

```{r}
controller$terminate()
```

Finally, use the process monitoring interface of your computing platform or operating system to verify that all `mirai` dispatchers and `crew` workers are terminated.

# Load testing

If the informal testing succeeded, we recommend you scale up testing to more ambitious scenarios. As one example, you can test that your workers can auto-scale and quickly churn through a large number of tasks.

```{r}
library(crew)
controller <- crew_controller_custom(
  seconds_idle = 2L,
  workers = 2L
)
controller$start()
# Push 100 tasks
for (index in seq_len(100L)) {
  name <- paste0("task_", index)
  controller$push(name = name, command = index, data = list(index = index))
  message(paste("push", name))
}
# Wait for the tasks to complete.
controller$wait()
# Wait for the workers to idle out and exit on their own.
crew_retry(
  ~all(controller$client$summary()$online == FALSE),
  seconds_interval = 1,
  seconds_timeout = 60
)
# Do the same for 100 more tasks.
for (index in (seq_len(100L) + 100L)) {
  name <- paste0("task_", index)
  controller$push(name = name, command = index, data = list(index = index))
  message(paste("push", name))
}
controller$wait()
crew_retry(
  ~all(controller$client$summary()$online == FALSE),
  seconds_interval = 1,
  seconds_timeout = 60
)
# Collect the results.
results <- NULL
while (!is.null(result <- controller$pop(scale = FALSE))) {
  if (!is.null(result)) {
    results <- dplyr::bind_rows(results, result)
  }
}
# Check the results
all(sort(unlist(results$result)) == seq_len(200L))
#> [1] TRUE
# View worker and task summaries.
controller$summary()
controller$schedule$summary()
controller$launcher$summary()
controller$schedule$summary()
# Terminate the controller.
controller$terminate()
# Now outside crew, verify that the mirai dispatcher
# and crew workers successfully terminated.
```

# Tips

* The [source code of the built-in local process launcher](https://github.com/wlandau/crew/blob/main/R/crew_launcher_local.R) is a helpful reference for writing your own launcher plugins.
* For efficiency, every launch can happen asynchronously. There is no need to wait for the worker to start. When it starts, it will connect back to the `mirai` dispatcher  over the local network and start accepting tasks.
* If you implement a [`terminate_worker()`](https://wlandau.github.io/crew/reference/crew_class_launcher_local.html#method-terminate-worker-) method, each worker termination may also happen asynchronously. In rare cases when you do not trust the platform to terminate the worker on the first request, you can use [`crew_retry()`](https://wlandau.github.io/crew/reference/crew_retry.html) to wait for the job to exit, but this may reduce efficiency.
