---
title: "Specification"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Specification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(crew)
```

This vignette is a specification of the classes in `crew`. Sub-classes are listed, but only the ones implemented in the `crew` package itself. Other sub-classes may be developed in packages that depend on `crew`.

# crew

A `crew` represents a group of workers. It contains:

* Its name.
* A `store` object with a data store to send and receive worker data.
* A list of worker objects to manage active workers
* A worker definition to launch new workers.

A crew is in charge of

* Launching new workers.
* Assigning jobs to workers.
* Polling workers.
* Receiving and processing data from workers.
* Terminating workers.

# worker

A `worker` object encapsulates a worker: a local process, cluster process, or cloud process that accepts and runs work that the crew dispatches. It contains:

* Its name.
* The `crew` object of the crew it belongs to.

## Sequential worker

Worker that runs in a simple R process.

## Future worker

Worker that runs in a `future`.

# store

A `store` object manages files on a file system that the workers and the crew both can access. The store object contains:

* Paths to directories and sub-directories.
* Methods to read, write, check the existence of, and delete worker files.

A store creates and manages one file per active worker. Files may exist locally or on the cloud, depending on the subclass. There is a 

## Local store

Subclass of a store based on local files.

## AWS S3 store

Subclass of a store based on AWS S3

## Google Cloud store

Subclass of a store based on Google Cloud Storage.

